Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-12-03 files/Overview.tex
overview

In the first chapter, the models of this dissertation is presented and a dimensionless variational form is derived. The model problems used in the dissertation is also presented.

In the second chapter, we look at the existence and uniqueness of solutions for a general vibration problem. All of the models in this dissertation is a special case of this general vibration problem. So a cantilever Timoshenko beam model is used as an example of the application of the theory as well as to explain the theory. A weak variational form of a cantilever Timoshenko beam model is derived. The authors provide four assumptions, and using these assumptions they prove using semi-group theory that the general vibration problem has a unique solution. For the example, we show that the four assumptions can be proven for the cantilever Timoshenko beam model.

Then we look at modal analysis. Again we use the cantilever Timoshenko beam model as an example. A trial solution is chosen and using this trial solution, two problems are obtained. The first is the eigenvalue problem and the second is an ordinary differential equation. These two problems are solved and substitution then shows that the trial solution is valid. This is similar to the separation of variables, but it avoids any division. We then discuss a general solution using the same idea. The idea of modal analysis is crucial to the dissertation. From modal analysis, we can compare the eigenvalues and eigenfunctions of the models, and if they compare well, then the solutions of the models will compare well. And since the formal series solution is a linear combination of Fourier Coefficients, we only need to consider the first few eigenvalues.

In the third chapter, two theoretical results are considered for the Finite Element Method. The first results is on the convergence of the Galerkin Approximation of a general vibration problem. In this section, the results of the article is discussed. The results are also given in Theorems with simple notation. We then look at a result form a textbook on the convergence of the eigenvalues and eigenfunctions when using the Finite Element Method. We discuss the results of the textbook, with some improvements to the notation as well as expanding some results to make it more clear.

>>>>

In the first chapter, the models of this dissertation are presented and a dimensionless variational form for each is derived. The model problems for this dissertation are then formulated.

In the second chapter, two results are investigated. This first is on the existence and uniqueness of solutions for vibration problems. An article is discussed that proves the existence and uniqueness of solutions for a general vibration problem using semi-group theory and assuming four assumptions. All the models in this dissertation is a special case of this general problem, so the theory can be explained and applied to an example. The weak variational form for a cantilever Timoshenko beam is derived. In this derivation, we also give the definitions of the spaces required. The theory is then applied to this example problem. We show that the four assumptions can be proven to be true for the cantilever Timoshenko beam model.

The next result is in Modal Analysis. Again the cantilever Timoshenko beam model is used to first explain the theory, before the general case is discussed. A trial solution is presented and substituted into the partial differential equations. This results into two problems, an eigenvalue problem and an ordinary differential equation. These two problems are solved and substitution can confirm then that the trial solutions are indeed solutions. The method obtains the same result as separation of variables, but avoids any division. Finally a formal series solution for the example is given. The general case is then discussed and follows the same idea. Modal analysis is crucial for this dissertation. For the comparisons later in this dissertation, a method is required to compare the different models. Modal analysis shows that is the eigenvalues and eigenfunctions of the models compare well, then the solutions of the models should compare well. And since the Fourier Coefficients gets small very fast, it is only required to look at the first few eigenvalues.

In the third chapter, two theoretical results for the Finite Element Method is discussed. The first result is on the convergence of the Galerkin Approximation. The general case contains various symbols rectangularly used in application, so an example of the cantilever Timoshenko beam model is used and the Galerkin Approximation is derived. Then the general case is discussed from an article. The results from the article are presented in a reduced format to avoid introducing too many new symbols. We also present the results in four new Theorems, to present the results of the article in a concise manner.

The second result is on the convergence of the eigenvalues and eigenfunctions when using the Finite Element Method. The results are presented in textbook. Here we use the notation of the dissertation to present the results. The results are also expanded in an attempt to better explain the theory, with a few new parameters also introduced.

In the fourth chapter, we investigate our simplest model, the Timoshenko beam theory. First is a discussion of an article that provides a method that allows us to find the exact eigenvalues and eigenfunctions of a Timoshenko beam model. A summary of this method is presented. It is then applied to two model problems for a cantilever Timoshenko beam and a free-free Timoshenko beam. For both model problems, the method is applied and examples of the eigenvalues and mode shapes are given. The eigenvalues and mode shape plots are obtained using a computer program, sketching the equation of motion and the using interval division to get the eigenvalues to a desired level of accuracy. This is in preparation for the comparison results in the results that follow.
